---
title: "MA585 PAPER"
author: "Xiru Lyu"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  pdf_document:
    highlight: haddock
    number_sections: yes
header: \usepackage{float}
geometry: margin=0.8in
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "h")

source("Analysis.R")
```

The paper talks about box office data.

\newpage


# Introduction


# Background

```{r fig.cap="Box Office Plot", fig.width=6, fig.height=2.5, fig.align="center"}
ggplot(train_p, aes(Time, Box)) + geom_line() +
  scale_x_date(date_labels = "%b-%Y", date_breaks = "3 months") + 
  xlab("") + ylab("Gross Box Office") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        plot.title = element_text(size = 14, hjust = 0.5)) + 
  ggtitle("Box Office Data")
```

# Modeling & Diagnostics

## Descriptive Analysis

```{r fig.width=8, fig.height=4, fig.cap="Data Decomposition"}
plot(z)
```

## Transformation to Stationary Process

```{r fig.width=6, fig.height = 3, fig.cap="Differenced Data", fig.align='center'}
plot(diff, main = "Differenced Box Office Data", ylab = "diff(1,12)")
```

## ARIMA Model

```{r fig.cap="ACF and PACF Plots for Training Data", fig.height=3.5}
par(mfrow=c(1,2))
acf(diff, lag.max = 48)
pacf(diff, lag.max = 48)
```

```{r fig.align='center', fig.cap="ARIMA subset selection"}
plot(fit_armasub)
```

**Model**     | **Full Model**                                | **AICc**
--------------| ----------------------------------------------|------------------
ARIMA(1,2,0)  | $X_{t}+0.396X_{t-1}=e_{t}$                    |`r fit_arima_1$aicc`
ARIMA(0,2,1)  | $X_{t}=e_{t}-0.999e_{t-1}$                    |`r fit_arima_2$aicc`
ARIMA(1,2,1)  | $X_{t}-0.011X_{t-1}=e_{t}-0.999e_{t-1}$       |`r fit_arima_3$aicc`
ARIMA(0,2,12) | $X_{t}=e_{t}+0.367e_{t-12}$                   |`r fit_arima_4$aicc`

Table: Comparision of Fitted ARIMA models

### Diagnostics

```{r fig.cap="Residual Analysis for ARIMA(3,2,12)"}
tsdiag(fit_arima_2)
```

```{r fig.cap="Normal Q-Q Plot for Fitted ARIMA Model", fig.height=4, fig.align='center'}
qqnorm(residuals(fit_arima_2), main = "Normal Q-Q Plot \nfor ARIMA(0,2,1)")
qqline(residuals(fit_arima_2))
```

## SARIMA model

**Model**                          |**AICc**
-----------------------------------|--------
SARIMA$(0,1,1)\times(0,1,1)_{12}$  |1032.6
SARIMA$(1,1,1)\times(0,1,1)_{12}$  |1034.8
SARIMA$(1,1,0)\times(0,1,1)_{12}$  |1054.3

Table: Comparison of Fitted SARIMA Models

### Diagnostics

```{r fig.cap="Residual Analysis for SARIMA(0,1,1)(0,1,1)[12]", fig.height=4}
tsdiag(fit_sarima_1)
```

```{r fig.cap="Residual Analysis for SARIMA(1,1,1)(0,1,1)[12]", fig.height=4}
tsdiag(fit_sarima_2)
```

```{r fig.cap="Comparison of Normal Q-Q Plots for Two Fitted SARIMA Models", fig.height=4}
par(mfrow=c(1,2))
qqnorm(residuals(fit_sarima_1), main = "Normal Q-Q Plot \nfor SARIMA(0,1,1)(0,1,1)[12]")
qqline(residuals(fit_sarima_1))

qqnorm(residuals(fit_sarima_2), main = "Normal Q-Q Plot \nfor SARIMA(1,1,1)(0,1,1)[12]")
qqline(residuals(fit_sarima_2))
```

# Forecast

**Criteria** | **ARIMA(0,2,1)**   | **SARIMA_1**         |**SARIMA_2**          | **Holt-Winters**
---------|----------------|------------------|------------------|---------
  RMSE   | `r ARIMArmse`  | `r SARIMArmse_1` | `r SARIMArmse_2` | `r HWrmse`
  MAPE   | `r ARIMAmapez` | `r SARIMAmapez_1`| `r SARIMAmapez_2`| `r HWmapez`
  MAE    | `r ARIMAmae`   | `r SARIMAmae_1`  | `r SARIMAmae_2`  | `r HWmae`
  
Table: Comparision of Forecast Accuracy


SARIMA_1: SARIMA$(0,1,1)\times(0,1,1)_{12}$

SARIMA_2: SARIMA$(1,1,1)\times(0,1,1)_{12}$

# Discussion

# Conclusion